How predictable is extinction?

### 2019-07-08 doc/napc2019_talk.tex
process for resubmit

  \begin{center}
    \begin{LARGE}
      Extinction is \alert{hard} to predict, but is \alert{important} to conservation decisions.
    \end{LARGE}
  \end{center}

@@

  \begin{itemize}[<+->]
    \item A taxon with a \alert{greater than average} global geographic range is likely to \alert{survive for longer} than a taxon with \alert{less than average} global geographic range.
    \item A taxon's global geographic range can change over time.
    \item What happens to extinction risk as a taxon changes geographic range? How is extinction risk impacted if that taxon's global geographic range has recently \alert{increased} or \alert{decreased}?
  \end{itemize}

@@

  \begin{itemize}%[<+->]
    \item Encoding the past
      \begin{itemize}
        \item Change in geographic range between current observation and previous observation.
        \item Average global temperature at time of previous observation (Mg/Ca elemental ratio).
        \item Age in millions of years at time of observation.
      \end{itemize}
    %\item \alert{Compare} models using WAIC/LOOIC.
    \item Explore model adequacy using posterior predictive distribution.
    \item Estimate out-of-sample predictive performance using \(k\)-fold cross-validation.
  \end{itemize}

@@

    \end{column}
    \begin{column}{0.6\textwidth}
      The area represents the probability of correct ranking of a random ``extant''-``extinct'' pair.

@@

@@

  \begin{itemize}
    \item \alert{The past matters\dots} 
      \begin{itemize}
        \item Our best supported model includes either our historical covariates or allows all effects to vary over time.
      \end{itemize}
    \item \alert{But not that much\dots}
      \begin{itemize}
        \item Models only average/fair expected out-of-sample performance.
      \end{itemize}
    \item Allowing effects to vary over time is probably preferable to historical covariates -- measures and accounts for variation which is important when predicting extinction in novel environments.
    \item Mechanisms behind changes to geographic range operate at sub-million year scales. Perhaps their effects are weak/masked at million (or greater) year scales.
  \end{itemize}

>>>>

  \begin{center}
    \begin{LARGE}
      Extinction is \alert{hard} to predict, \\but is \alert{important} to conservation decisions.
    \end{LARGE}
  \end{center}

@@

  \begin{Large}
    \begin{itemize}[<+->]
      \item A taxon with a \alert{greater than average} global geographic range is likely to \alert{survive for longer} than a taxon with \alert{less than average} global geographic range.
      \item A taxon's global geographic range can change over time.
      \item What happens to extinction risk as a taxon changes geographic range? How is extinction risk impacted if that taxon's global geographic range has recently \alert{increased} or \alert{decreased}?
    \end{itemize}
  \end{Large}

@@

  \begin{Large}
    \begin{itemize}%[<+->]
      \item Encoding the past
        \begin{itemize}
            \begin{large}
            \item Change in geographic range between current observation and previous observation.
            \item Average global temperature at time of previous observation (Mg/Ca elemental ratio).
            \item Age in millions of years at time of observation.
            \end{large}
        \end{itemize}
      \item Explore model adequacy using posterior predictive distribution.
      \item Estimate out-of-sample predictive performance using \(k\)-fold cross-validation.
    \end{itemize}
  \end{Large}

@@

    \end{column}
    \begin{column}{0.6\textwidth}
      \begin{large}

      The area represents the probability of correct ranking of a random ``extant''-``extinct'' pair.

@@

      \end{large}

@@

  \begin{large}
    \begin{itemize}
      \item \alert{The past matters\dots} 
        \begin{itemize}
          \item Our best supported model includes either our historical covariates or allows all effects to vary over time.
        \end{itemize}
      \item \alert{But extinction is still hard to predict}
        \begin{itemize}
          \item Models only average/fair expected out-of-sample performance.
        \end{itemize}
      \item Allowing effects to vary over time is probably preferable to historical covariates -- measures and accounts for variation which is important when predicting extinction in novel environments.
      \item Mechanisms behind changes to geographic range operate at sub-million year scales. Perhaps their effects are weak/masked at million (or greater) year scales.
    \end{itemize}
  \end{large}
